### 2025-01-22 Thesis/Latex/Chapter_3.tex
Updated thesis template with content for chapters

\chapter{Neural Networks and Machine Translation}
\pagestyle{fancy}\lhead{\textbf \footnotesize\it{Neural Networks and Machine Translation}}
\pagestyle{fancy}\chead{} \pagestyle{fancy}\rhead{}
\pagestyle{fancy}\lfoot{\textbf {\small\it{Univ-Mascara/Computer Science: 2024}}} 
\pagestyle{fancy}\cfoot{} \pagestyle{fancy}\rfoot{\thepage}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Overview}\label{start3}
In this chapter, we embark on an in-depth exploration of neural networks, pivotal to understanding the intricate workings of machine translation. 
Beginning with a thorough examination of the foundational feed-forward networks (FFNs), recurrent neural networks (RNNs), long short-term memory networks (LSTMs), and gated recurrent units (GRUs), we lay the groundwork for comprehending the neural machine translation (NMT) process.

We delve into the nuances of training NMT models, and the decoding phase, where trained models generate translations from input sequences. 
Finally, we conducted a comprehensive survey of recent advancements and methodologies in neural machine translation research, with a focus on exploring related works specifically dedicated to the Arabic language context.

\section{Neural Networks}
Neural networks are the foundational computational mechanisms driving Neural Machine Translation, employing complex algorithms to mimic human language understanding and generation. 
At the core of these networks is the node, or processing unit, which functions similarly to neurons in the human brain. 
Each node is designed to receive inputs—typically a vector of real-valued numbers—process these inputs through a series of mathematical operations, including weighted sums and activation functions, and then produce an output that can be passed on to subsequent layers or nodes in the network.
The following subsections will delve into the different types of neural networks, beginning with the simplest form: the feed-forward neural network.

\subsection{Feed-Forward Neural Networks}
A feed-forward neural network is a basic type of neural network where information moves in one direction from the input to the output layer without any loops or cycles back to previous layers. 
Outputs from each node are forwarded upwards to the subsequent layer without any feedback to lower levels. 
An illustration, referred to as Figure~\ref{fig:ffn}, shows a FFN consisting of two layers. 
These networks are composed of three distinct types of nodes: input, hidden, and output. 
The input nodes receive individual scalar values, and notably, among these input nodes is a constant bias node, labeled $x_0$, which is always set to the value of 1. 
In the hidden layer, nodes calculate a weighted sum of their inputs and then apply a nonlinear function to this sum to determine their output.
\begin{figure}[h]
	\centerline{\includegraphics[width=0.7\linewidth]{Figures/FFN}}
	\caption{A feed-forward network consisting of two layers includes an input layer $x$, a hidden layer $h$, and an output layer $y$.}
	\label{fig:ffn}
\end{figure}
In a standard neural network structure, every layer is comprehensively interconnected, allowing nodes to receive inputs from all nodes in the preceding layer. 
These networks are characterized by their depth, with numerous layers contributing to their complexity. 
Hidden nodes within these layers act as automatic feature detectors, eliminating the need for manual feature identification by learning to recognize relevant patterns in the input data through training. 
Each hidden node is defined by its parameters, including a weight vector and a bias term. 
The parameters for the entire hidden layer are aggregated into a weight matrix $U$, where each weight $u_{jk}$ in $U$ corresponds to the connection strength from the $k^{th}$ input node $x_k$ to the $j^{th}$ hidden node $h_j$, and a bias vector that applies to the entire layer.

A notable feature of neural networks is the incorporation of non-linear activation functions, with the rectified linear unit ($ReLU$) standing out for its simplicity and widespread usage. 
$ReLU$ is particularly efficient to compute, as depicted in Figure~\ref{fig:relu} (a). 
This function returns the value of $z$ when $z$ is greater than 0 and returns $0$ when $z$ is less or equal to 0.
\begin{figure}[h]
	\centering
	\includegraphics[width=0.7\linewidth]{Figures/Relu}
	\caption{Common activation functions utilized in neural networks.}
	\label{fig:relu}
\end{figure}
The $sigmoid$ (or logistic) function, another frequently utilized activation function, has been illustrated in Figure~\ref{fig:relu} (b).
An activation function similar to the sigmoid but more widely employed is the hyperbolic tangent ($tanh$), depicted in Figure~\ref{fig:relu} (c).

The computation of the hidden layer in the basic feed-forward network can be performed efficiently using straightforward matrix operations, involving three main steps:
\begin{itemize}
	\item The multiplication of the weight matrix by the input vector $x$,
	\item Addition of the bias vector, and
	\item Application of the activation function $f$, such as $ReLU$, $sigmoid$, or $tanh$.
\end{itemize}
Hence, the neural network depicted in Figure~\ref{fig:ffn} can be expressed mathematically as follows: 
\begin{itemize}
	\item A set of input nodes represented by the vector $x = (x_1, x_2, x_3, . . . , x_n)^T$;
	\item A set of hidden nodes represented by the vector $h = (h_1, h_2, h_3, . . . , h_l)^T$;
	\item A set of output nodes represented by the vector $y = (y_1, y_2, y_3, . . . , y_m)^T$;
	\item A weight matrix connecting input nodes to hidden nodes denoted by $U = u_{jk}$;
	\item A weight matrix connecting hidden nodes to output nodes denoted by $W = w_{ij}$.
\end{itemize}
The computation of the hidden layer output, represented by the vector $h$, is performed according to Equation~\ref{eqn:hj}, incorporating the activation function $f$.
\begin{equation}
	h_j = f(\sum_{k}u_{jk}x_k)
	\label{eqn:hj}
\end{equation}
The value $h$ obtained serves as a representation of the input. 
Subsequently, the output layer's function is to utilize this revised representation $h$ and calculate a conclusive output, as described in Equation~\ref{eqn:yi}.
\begin{equation}
	y_i = \sum_{j}w_{ij}h_j
	\label{eqn:yi}
\end{equation}
In numerous instances, this output may initially be a real-valued number; however, it is often transformed into a probability distribution using a $softmax$ function. For a vector $y$ with a dimensionality of $d$, the $softmax$ function is defined as shown in Equation~\ref{eqn:softmax_yi}, where $1 \leq i \leq d$.
\begin{equation}
	softmax(y_i) =\frac{e^{y_i}}{\sum_{j}e^{y_j}}
	\label{eqn:softmax_yi}
\end{equation}
The $softmax$ function operates on a vector $y = [y_1, y_2, y_3, . . . , y_m]$ of arbitrary values, transforming them into a probability distribution where each value falls within the range of (0, 1) and the sum of all values equals one.

\subsection{Recurrent Neural Networks}
A neural network containing a cycle inside its network connections is called a recurrent neural network (RNN). 
Its preceding outputs directly or indirectly influence the value of a node. 
Figure~\ref{fig:rnn} illustrates the structure of a simple RNN based on \cite{elman90}. 
Similar to conventional FFNs, the values for a layer of hidden nodes are calculated by multiplying an input vector representing the current input, $x$, by a weight matrix and then passing the result through a non-linear activation function. 
The associated output, $y$, is then determined using the hidden layer, which comprises the hidden nodes. 
The context layer is where it differs from a FFN the most. 
It keeps the previous values and sends them to the appropriate nodes in the hidden layer. 
This layer uses the hidden layer’s value from the previous time step as input to the computation at the hidden layer.
\begin{figure}[h]
	\centering
	\includegraphics[width=0.7\linewidth]{Figures/RNN}
	\caption{A basic recurrent neural network.}
	\label{fig:rnn}
\end{figure}

\subsubsection{Stacked Recurrent Neural Networks} 
Stacked recurrent neural networks (RNNs) consist of multiple layers, where the output of one layer serves as the input to the next layer. 
Information flows from lower layers to higher layers, and the final output is generated by the top layer. 
Stacked RNNs have shown improved performance in neural MT compared to single-layer networks \cite{bahdanau15}. 
This improvement can be attributed to the network's ability to create representations at different levels of abstraction across the layers. 
However, the optimal number of stacked RNNs depends on the availability of training data. 
While abundant data can lead to better generalization, limited data may result in overfitting \cite{lample18}. 
Moreover, increasing the number of stacked layers escalates the training costs.

\subsubsection{Bidirectional Recurrent Neural Networks}
In recurrent networks, the hidden state at any given time encapsulates all information about the sequence up to that point, serving as the left context for the current input. 
In neural machine translation (NMT), having access to the entire input sequence simultaneously suggests the importance of utilizing the context to the right of the input. 
Training an RNN in reverse on the input sequence is one approach to retain this information. 
Combining the forward and backward networks results in a bidirectional RNN, where two independent RNNs process input from start to finish and finish to start, respectively \cite{schuster97}. 
The outputs of both networks are then merged to create a unified representation incorporating both left and right input contexts at each time step. 
Methods such as concatenation, element-wise operations, or averaging can be employed to combine the outputs of the forward and backward passes. 
This ensures that the information on both sides of the current input is captured in the output at each time step.
Figure~\ref{fig:rnn-brnn} illustrates the conventional RNN structure alongside the BRNN architecture over a timeline. In the RNN, only a forward hidden layer is present, whereas in the BRNN, both forward and backward hidden layers are depicted.

\begin{figure}[h]
	\centering
	\includegraphics[width=0.9\linewidth]{Figures/RNN-BRNN}
	\caption{Architecture of RNN and BRNN shown over a period of time.}
	\label{fig:rnn-brnn}
\end{figure}

\subsection{Long Short-Term Memory}
One of the main limitations of RNNs in machine translation is their difficulty in handling long-distance dependencies. Distant words play a crucial role in translation tasks, as demonstrated in the following example:
\begin{quotation}
	The lengthy \textit{sentence} composed of numerous words spanning multiple lines still \textit{poses} significant challenges for MT.
\end{quotation}
In this example, the verb \textit{poses} depends on the subject \textit{sentence}, which is separated by a long subordinate clause.

While RNNs can access preceding sequences, the knowledge retained in hidden states tends to be relatively localized. 
Consequently, more complex network structures have been developed to address the difficulty of maintaining context over extended periods. 
It is essential for the network to discard irrelevant information while retaining data crucial for future decision-making. 
Long Short-Term Memory (LSTM) and Gated Recurrent Units (GRUs) are the predominant approaches employed to achieve this objective.

LSTM networks, introduced by \cite{hochreiter97}, address the challenge of managing context by dividing it into two sub-problems: removing unnecessary information and incorporating data crucial for future decisions. 
Unlike incorporating a fixed strategy into the architecture, the key lies in learning to handle context dynamically. 
LSTMs accomplish this by expanding the architecture with an explicit context layer and employing specialized neural units with gates to control the flow of information within the network layers. 
These gates utilize additional weights to sequentially process input, previous hidden layers, and previous context layers.

\subsection{Gated Recurrent Units}
GRU, akin to LSTM but with fewer parameters, offers the advantage of reduced training costs by requiring fewer parameters. 
This reduction in parameters is achieved by eliminating the need for a separate context vector and reducing the number of gates, thereby alleviating the computational burden compared to LSTM. 
Both GRU and LSTM employ sigmoid functions in their gates to either allow information with values close to one to pass through or block information with values close to zero.

In contrast to simple FFNs, LSTMs and GRUs feature more complex neural nodes, as depicted in Figure~\ref{fig:lstm-vs-gru}, with inputs and outputs connected to each node type. 
The additional complexity within LSTM and GRU nodes is primarily contained within the nodes themselves. 
The only external complexity introduced by LSTM, beyond the basic recurrent node, is the availability of the other context vector as an input and output. 
Conversely, GRU nodes exhibit similar input and output architecture to the simple recurrent node.
\begin{figure}[h]
	\centering
	\includegraphics[width=0.7\linewidth]{Figures/LSTM-vs-GRU}
	\caption{Diagram illustrating the network structure of LSTM in the first panel (a) and GRU in the second panel (b)}
	\label{fig:lstm-vs-gru}
\end{figure}

\section{Neural Machine Translation}
The standard architecture for Neural Machine Translation (NMT) is the encoder-decoder network. 
This architecture operates on the fundamental principle of employing an encoder network to convert a sequence of words from the source language sentence into a contextual representation. 
Subsequently, a decoder utilizes this representation to generate an output sequence, essentially providing a plausible translation into the target language. 
Figure~\ref{fig:encoder-decoder} illustrates the encoder-decoder architecture at its most abstract level, comprising three key components: an encoder, context, and decoder. 
The encoder takes in a source language sentence as an input sequence of words, $x_1, x_2, ..., x_m$, and produces a corresponding sequence of contextualized representations known as a context vector. 
This context vector captures the essence of the input and serves as input to the decoder. 
Finally, the decoder utilizes the context vector to generate the most probable translation as a sequence of words, $y_1, y_2, ..., y_n$. 
Encoders and decoders can be implemented using either RNNs or Transformers.
\begin{figure}[h]
	\centering
	\includegraphics[width=0.7\linewidth]{Figures/Encoder-Decoder}
	\caption{Basic Encoder-Decoder architecture}
	\label{fig:encoder-decoder}
\end{figure}

Figure 2.7 illustrates a simplified version of the RNN-based encoder-decoder architecture. 
The encoder processes the input sequence $x$ with the objective of generating a representation of the input. 
This representation is captured in the final hidden state of the encoder, represented as $h^e_n$. 
Subsequently, this context representation, denoted as $c$, is transferred to the decoder. 
Upon receiving this state, the decoder utilizes it to initialize its initial hidden state. 
The first decoder cell employs $c$ as its initial hidden state, $h^d_0$, and proceeds to generate a sequence of outputs iteratively, one element at a time, until an end-of-word marker, </s>, is produced. 
Consequently, each hidden state is dependent on the preceding hidden state and the output generated in the preceding state. 
The embedding layer is composed of word embeddings, which adhere to the notion that semantically related words in similar contexts should possess similar representations. 
Ultimately, the output $y$ at each time step involves a $softmax$ computation over the vocabulary, $V$. The most probable output at each time step can be determined by computing the $argmax$ over the $softmax$ output as per Equation~\ref{eqn:yt}. 
Although Figure~\ref{fig:rnn-encoder-decoder} illustrates a single network layer, stacked and bi-directional networks are typically employed for both the encoder and decoder in practical implementations.
\begin{equation}
	\hat{y}_t = argmax_{w \in V}P(w|x, y_1, . . . , y_{t-1})
	\label{eqn:yt}
\end{equation}
\begin{figure}[]
	\centering
	\includegraphics[width=1.0\linewidth]{Figures/RNN-Encoder-Decoder}
	\caption{The basic RNN-based encoder-decoder architecture.}
	\label{fig:rnn-encoder-decoder}
\end{figure}

A significant drawback of this architecture lies in its inability to evenly represent information from the beginning of a sentence, particularly evident in lengthy sentences. 
To address this issue, the attention mechanism emerges as a solution. 
It enables the decoder to access information from all the hidden states of the encoder, rather than solely relying on the last hidden state. 
The concept behind the attention mechanism is to create a singular context vector, denoted as $c$, by computing a weighted sum of all the encoder's hidden states. 
These weights concentrate on a specific segment of the source text pertinent to the token generated by the decoder. 
Notably, the context vector produced by the attention mechanism is dynamic, varying for each decoded token. 
By introducing the attention mechanism, the static context vector is replaced with $c_i$ a dynamically derived counterpart from the encoder's hidden states at each decoding step $i$, ensuring comprehensive consideration of all encoder states.

\subsection{Training Neural Machine Translation Models}
Neural Machine Translation (NMT) operates within a supervised machine learning framework, where the correct output, denoted as $y$, is known for each input observation, $x$. 
The system's output, represented as $\hat{y}$, serves as an estimate of the actual $y$. 
During the training phase, the objective is to adjust the parameters within each layer so that $\hat{y}$ closely resembles $y$ for every training instance.
To accomplish this, a loss function is employed to quantify the disparity between the system's output and the actual output. 
The optimization process, typically facilitated by the gradient descent algorithm, seeks to minimize this loss function by adjusting the parameters. 
Gradient descent utilizes the gradient of the loss function, computed as the partial derivative of the loss function with respect to each parameter, to iteratively update the model's parameters. 
However, in the realm of neural networks, characterized by numerous layers and millions of parameters, computing the partial derivative of a weight in one layer concerning the loss associated with another layer necessitates employing techniques such as error back-propagation or reverse differentiation. 
These methodologies are crucial for efficiently navigating the complex landscape of neural network optimization.

Moreover, NMT models undergo end-to-end training, where each training instance comprises a pair of sentences, one in the source language and the other in the target language. These sentence pairs are concatenated with a designated separator token, <s>, to form the training data. Typically, this training data is sourced from established datasets containing aligned pairs of sentences, known as parallel corpora. Optimization within the NMT framework is characterized by a non-convex problem landscape.
Nevertheless, there exist several best practices for effectively training NMT models. For instance, it is advisable to initialize the model weights with small random numbers and to employ random seeds for reproducibility. Additionally, normalizing the input values to have zero mean and unit variance can be beneficial in enhancing training stability and convergence.

NMT model training involves multiple epochs, representing complete iterations over the training data. Typically, as training progresses, the error on the training set steadily decreases. However, a key challenge arises when the model starts to overfit, meaning it memorizes the training data excessively without generalizing well to unseen examples. To diagnose overfitting, a separate set of examples known as the development (or validation) set is used, which is not involved in the training process. Monitoring the error on this development set over the course of training reveals a point where the error begins to increase, indicating overfitting.
In theory, training should halt when the minimum error on the development set is reached. However, in practice, this is more nuanced for NMT models due to the non-deterministic nature of training and the lack of clear convergence or overfitting points, particularly with sizable datasets. While some studies suggest stopping criteria based on approximate training durations or a fixed number of epochs, most NMT research does not specify precise stopping conditions. To mitigate overfitting, various regularization techniques are employed, such as dropout, which randomly disables some nodes and connections during training.

In addition to regularization, hyperparameter tuning plays a crucial role in NMT architecture design. These hyperparameters, which include the learning rate, mini-batch size, number of layers, hidden nodes per layer, and activation functions, are chosen by the architect to optimize model performance.
\subsection{Decoding}
The decoding (inference) algorithm employed to generate translations (as illustrated in the Figure~\ref{fig:rnn-encoder-decoder}) faces a challenge. 
Opting for the single most probable word at each step implies a 1-best greedy search, where a greedy algorithm makes locally optimal decisions. 
However, there are instances where following a sequence of words leads to the realization that an earlier mistake was made. 
In such cases, the best sequence may initially comprise less probable words that are refined by subsequent words in the context of the entire output. 
For instance, when translating an idiomatic expression, the initial words chosen may seem unusual (e.g., "\textit{take your hat off}" for "\textit{respect, admire, or congratulate someone}").

>>>>

\chapter{RAG (Retrieval-Augmented Generation)}
\pagestyle{fancy}\lhead{\textbf \footnotesize\it{Enhancing Legal Information Access with Retrieval Augmented LLMs for Juridical Data}}
\pagestyle{fancy}\chead{} \pagestyle{fancy}\rhead{}
\pagestyle{fancy}\lfoot{\textbf {\small\it{Univ-Mascara/Computer Science: 2025}}} 
\pagestyle{fancy}\cfoot{} \pagestyle{fancy}\rfoot{\thepage}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{ Introduction}\label{start2}

Large language models (LLMs) have seen significant advancements but face challenges, particularly in tasks that demand extensive knowledge or deal with queries beyond their training data. These limitations often result in inaccuracies or “hallucinations.” To address this, Retrieval-Augmented Generation (RAG) supplements LLMs by retrieving relevant document chunks from external knowledge sources based on semantic similarity. By doing so, RAG helps reduce factual errors, allowing LLMs to produce more accurate content. This has led to the widespread adoption of RAG, particularly in chatbots and other real-world applications, making it a crucial technology in advancing the capabilities of LLMs.
\section{Fundamentals of Retrieval-Augmented Generation}
RAG represents a significant advancement in the capabilities of large language models (LLMs) by integrating external knowledge retrieval with the generation of text. Understanding the individual components of retrieval and generation is essential to appreciate how their synergy improves the overall performance of RAG systems.
\subsection{Definition of RAG}
Retrieval-Augmented Generation (RAG) is a technique that leverages the strengths of pre-trained large language models (LLMs) and external data sources. By merging the generative abilities of LLMs like GPT-3 or GPT-4 with the accuracy of specialized data search mechanisms, RAG systems can generate more sophisticated and contextually relevant responses\cite{selvaraj2024}.
\subsection{Historical Development of RAG}
The foundations of Retrieval-Augmented Generation (RAG) can be traced back to traditional information retrieval (IR) systems,  Early models primarily relied on keyword matching and simple ranking mechanisms to retrieve relevant documents from databases, establishing a basic framework for information access. The landscape began to shift with the rise of neural networks in the 2010s. Notable innovations like Word2Vec and Transformer-based architectures enabled retrieval methods to incorporate deeper semantic understanding, paving the way for more sophisticated document retrieval processes.
A significant advancement occurred with the introduction of dense passage retrieval (DPR) in 2020, which utilized bi-encoder architectures to map both queries and documents into dense vector spaces. 
The integration of these advanced retrieval techniques with large language models (LLMs) became a focal point as models like GPT-3 gained prominence. Researchers explored how LLMs could be augmented with retrieval capabilities, leading to the development of RAG systems that leverage the strengths of both retrieval and generative capabilities\cite{gao2024retrieval}.
\subsection{ Differences Between RAG and Fine-Tuning}
The enhancement of large language models (LLMs) has gained significant interest due to their increasing use. Among the various optimization strategies for LLMs, Retrieval-Augmented Generation (RAG) is often compared to fine-tuning (FT) and prompt engineering. Each method possesses unique attributes,
\begin{enumerate}
	\item Methodological Characteristics :\\
	RAG is compared to providing a tailored textbook for information retrieval, making it suitable for precise information retrieval tasks. It excels in dynamic environments with real-time knowledge updates. \\
	Fine-tuning is likened to a student internalizing knowledge, making it more static and suitable for replicating specific structures, styles, or formats.
	\item External Knowledge and Adaptation: \\
	RAG relies on external knowledge sources and allows for high interpretability but may involve higher latency and ethical considerations regarding data retrieval. \\
	Fine-tuning requires retraining for updates and involves significant computational resources for dataset preparation and training. It enables deep customization of the model’s behavior but may struggle with unfamiliar data.
	\item Performance: \\ RAG consistently outperforms unsupervised fine-tuning in knowledge-intensive tasks, particularly for both previously encountered and new knowledge.
	LLMs often struggle to learn new factual information through unsupervised fine-tuning.
	\item Use Cases and Combination: \\
	The choice between RAG and fine-tuning depends on specific needs for data dynamics, customization, and computational capabilities. They are not mutually exclusive; their combined use may yield optimal performance and often requires multiple iterations to refine the results \cite{gao2024retrieval}.

\end{enumerate}
\section{Types of RAG Systems}
The RAG research field is continuously evolving, with three main stages: Naive RAG, Advanced RAG, and Modular RAG.
\subsection{Naive RAG}
Naive RAG, a foundational approach in Retrieval-Augmented Generation, operates on a straightforward "Retrieve-Read-Generate" paradigm. This method involves three primary steps:

\begin{itemize}
	\item Indexing: Raw data is cleaned, extracted, and converted into a uniform text format. It's then segmented into smaller chunks and encoded into vector representations using an embedding model. These vectors are stored in a vector database for efficient similarity searches.
	\item Retrieval: When a user query is received, it's encoded into a vector and compared to the indexed chunks. The top K most similar chunks are retrieved and included in the prompt for the language model.
	\item Generation: The query and retrieved chunks are combined into a prompt, which is fed to a large language model. The model generates a response based on the provided context and its internal knowledge.
\end{itemize}
While Naive RAG offers a basic framework, it faces several challenges in 
\begin{itemize}
	\item Retrieval Issues: The retrieval process can be imprecise, leading to the selection of irrelevant or missing information.
	\item Generation Challenges: The language model may hallucinate, generating content not supported by the retrieved context. It may also produce irrelevant, toxic, or biased outputs.
	\item Augmentation Difficulties: Integrating retrieved information into the generation process can be challenging, leading to disjointed or redundant responses.
\end{itemize}
To address these limitations, more sophisticated RAG techniques have emerged, which we will explore in the next section.
\subsection{Advanced RAG}
Taking aim at the shortcomings of Naive RAG, Advanced RAG introduces specific improvements to enhance retrieval quality. This approach utilizes pre-retrieval and post-retrieval strategies.
\subsubsection{Pre-retrieval Strategies:}
\begin{itemize}
	\item Enhanced Indexing: Advanced RAG tackles indexing issues through a sliding window approach, finer segmentation of data, and inclusion of metadata. Additionally, it optimizes the retrieval process by employing various methods.
	\item Query Optimization: This stage focuses on refining the user's initial query to make it clearer and more suitable for retrieval. Techniques like query rewriting, transformation, and expansion are commonly used.
\end{itemize}
\subsubsection{Post-Retrieval Strategies:}
\begin{itemize}
	\item Re-ranking Chunks: After relevant information is retrieved, Advanced RAG prioritizes the most relevant content by re-ranking the retrieved chunks and placing them strategically within the prompt.
	\item Context Compression: To avoid overwhelming the LLM with too much information, post-retrieval efforts focus on selecting the most essential parts of the retrieved context, highlighting critical sections, and compressing the data to be processed.
\end{itemize}
By addressing indexing issues and refining the query and retrieved information, Advanced RAG aims to improve the overall accuracy and relevance of the generated response.
\subsection{Modular RAG} 
Modular RAG represents the latest evolution in RAG, offering greater adaptability. it introduces specialized modules and innovative patterns to enhance retrieval and processing capabilities.
\subsubsection{New Modules}
\begin{itemize}
	\item Search Module: Adapts to specific scenarios by leveraging LLM-generated code and query languages to search across various data sources.
	\item RAG Fusion: Employs a multi-query strategy to expand user queries, uncover both explicit and implicit knowledge, and improve retrieval results.
	\item Memory Module: Leverages the LLM's memory to guide retrieval and create an unbounded memory pool, aligning the text more closely with data distribution.
	\item Routing Module: Navigates through diverse data sources, selecting the optimal pathway for a query based on its specific needs.
	\item Predict Module: Reduces redundancy and noise by generating relevant context directly through the LLM.
	\item Task Adapter Module: Tailors RAG to various downstream tasks by automating prompt retrieval and creating task-specific retrievers.
\end{itemize}
\subsubsection{New Patterns} 
\begin{itemize}
	\item Flexible Module Arrangement: Modular RAG allows for the substitution and reconfiguration of modules to address specific challenges, surpassing the fixed structures of previous RAG paradigms.
	\item Innovative Retrieval Strategies: Techniques like Rewrite-Retrieve-Read, Generate-Read, and Recite-Read leverage the LLM's capabilities to refine queries, generate content, and retrieve information from model weights.
	\item Hybrid Retrieval: Combines keyword, semantic, and vector searches to cater to diverse queries, improving retrieval relevance.
	\item Dynamic Module Interaction: Frameworks like Demonstrate-Search-Predict and ITERRETGEN demonstrate the dynamic use of module outputs to enhance each other's functionality.
	\item Adaptive Retrieval: Techniques like FLARE and Self-RAG evaluate the necessity of retrieval based on different scenarios, allowing for a more flexible and efficient approach.

\end{itemize}
\section{Core Components of RAG}
The core components of Retrieval-Augmented Generation (RAG) systems consist of a retrieval mechanism, a generation process, and augmentation techniques. These elements work together to enhance the model's ability to access relevant external information, generate coherent and contextually appropriate responses, and improve overall performance in knowledge-intensive tasks.
\subsection{ Retrieval Mechanism}
RAG systems combine parametric memory (a pre-trained language model) with non-parametric memory (a retrieval mechanism). The retrieval mechanism allows RAG models to access external information sources (e.g., Wikipedia), and this process is central to improving the model's ability to generate factual and accurate outputs \cite{lewis2020retrieval}.
\begin{itemize}
	\item Traditional Techniques: \\
	Classical retrieval methods include \textbf{TF-IDF and BM25}, which rely on sparse vector representations based on term frequencies. These methods use exact keyword matching, making them limited in semantic understanding.
	\item Modern Retrieval Approaches:\\
	\textbf{ Dense Passage Retrieval (DPR): }This approach utilizes dense vector representations learned by neural models like BERT to encode both the query and document. It allows for a more semantic understanding of the content, making retrieval more effective. DPR, for example, computes the similarity between the query and documents using Maximum Inner Product Search (MIPS), which finds the closest passages based on the dense vector space.
	\item Implementation in RAG:\\ In RAG, dense retrieval methods are often paired with techniques like \textbf{Maximum Inner Product Search} (MIPS), which efficiently matches query and document embeddings. This enables the system to return the most relevant documents for subsequent generation \cite{karpukhin2020dense}.

\end{itemize}
\subsection{Generation Process}
The generation in RAG occurs through a combination of retrieved passages and the input query.\\
Large Language Models (LLMs) like BART or T5 are utilized for this generation, leveraging their advanced capabilities in understanding and producing human-like text. The retrieval component supplies external context, which the generator conditions on to formulate a response.This context is crucial, as it enriches the LLM's understanding, enabling it to incorporate real-time, relevant information. Moreover, the generation process benefits from the LLM's ability to synthesize information, allowing it to create responses that are not only coherent but also informed by the latest data, thus improving accuracy and relevance in knowledge-intensive tasks \cite{lewis2020retrieval}. \\
Two models have been proposed in RAG:
\begin{itemize}
	\item \textbf{RAG-Sequence} uses the same document to generate the entire sequence of output. It marginalizes over the top K retrieved documents to produce a final answer.
	\item \textbf{RAG-Token} allows different tokens in the output sequence to be generated based on different documents, making it more flexible when combining information from various
\end{itemize}
\subsection{ Augmentation Techniques}
The augmentation of the retrieval process in Retrieval-Augmented Generation (RAG) systems focuses on improving how queries are refined and how relevant information is retrieved for downstream generation tasks. Key augmentation methods include.
\begin{itemize}
	\item \textbf{Query Augmentation:} In traditional retrieval pipelines, user queries are often under-specified or ambiguous, leading to poor retrieval performance. Query augmentation involves dynamically rewriting the user's query to better match the documents in the knowledge base. This can be done by leveraging large language models (LLMs) to generate tailored queries or synthetic questions and answers (QAs) that better align with the search objective​
	\item \textbf{Synthetic QA Generation:} Instead of using raw document chunks, retrieval is enhanced by generating and embedding synthetic QA pairs from documents. This helps to capture the semantic essence of long texts more effectively, reducing noise and improving retrieval precision. These synthetic QAs can be used to rewrite the user query, making it more specific to the task at hand \cite{mombaerts2024meta}.
\end{itemize}
\begin{figure}
	\centering
	\includegraphics[width=0.9\linewidth]{Figures/rag_architectur.png}
	\caption{Rag architecture}
	\label{rag_architectur.png}
\end{figure}
\section{Training and Fine-Tuning RAG Models}
Training and fine-tuning are essential steps in enhancing the performance of RAG models by optimizing both the retriever and generator components leading to more coherent, relevant, and accurate responses.
\subsection{Training the Retriever}
A powerful technique for improving retrieval relevance involves training the retriever model using \textbf{contrastive learning}. This approach aims to maximize the similarity between query and relevant document embeddings while minimizing the similarity between query and irrelevant document embeddings. By training the model on numerous positive (relevant) and negative (irrelevant) pairs, the retriever learns to effectively distinguish between relevant and irrelevant information\cite{sbert2024}.
\section{Task and Evaluation}
The rapid development and increasing use of Retrieval-Augmented Generation (RAG) models in NLP have made their evaluation a critical area of research within the LLM community. This section covers the primary downstream tasks associated with RAG, the datasets used, and the methods for evaluating RAG systems \cite{zhou2020trustworthiness}.
\subsection{Factuality}
\begin{itemize}
	\item Fact-Checking Benchmarks: Using datasets like FEVER and SQuAD to evaluate the model's ability to identify and correct factual errors.
	\item Adversarial Testing: Creating adversarial examples to test the model's robustness against misleading information.
	\item Contextual Understanding: Assessing the model's ability to understand the context of a query and provide accurate answers.
\end{itemize}
\subsection{ Robustness}
\begin{itemize}
	\item Noise Injection: Introducing noise into the retrieved documents to test the model's ability to handle imperfect information.
	\item Adversarial Attacks: Evaluating the model's resilience to attacks that aim to manipulate its outputs.
	\item Domain Adaptation: Testing the model's ability to adapt to new domains and data distributions.
\end{itemize}
\subsection{Fairness}
\begin{itemize}

	\item Bias Detection: Identifying and quantifying biases in the training data and model outputs.
	\item Fairness Metrics: Using metrics like demographic parity and equalized odds to evaluate the model's fairness.
	\item Mitigation Techniques: Implementing techniques like debiasing and fairness constraints to mitigate biases.
\end{itemize}
\subsection{Objective Metrics:}
\begin{itemize}
	\item Accuracy: Precision, recall, and F1-score for evaluating the correctness of the generated responses.
	\item Consistency: Measuring the consistency of the model's outputs across different queries and contexts.
	\item Coherence: Assessing the coherence and fluency of the generated text.
\end{itemize}
\subsection{Subjective Metrics}
\begin{itemize}
	\item Human Evaluation: Using human raters to evaluate the quality of the generated responses.
	\item User Studies: Conducting user studies to gather feedback on the user experience
\end{itemize}
\section{Limitations}
While RAG has gained significant traction across diverse applications, it still faces certain limitations in terms of effectiveness and efficiency\cite{zhao2024retrieval}.
\subsection{Noisy Retrieval Results}
\begin{itemize}
	\item Retrieval Quality: The quality of retrieved information can be affected by factors like indexing techniques, query formulation, and the underlying dataset.
	\item Hallucinations: Noisy or irrelevant information can lead to the generation of hallucinated or factually incorrect responses.
	\item Contextual Understanding: The LLM may struggle to understand the context of the retrieved information, especially if it's poorly formatted or contains inconsistencies.
\end{itemize}
\subsection{Extra Overhead}
\begin{itemize}
	\item Computational Cost: Retrieval and processing additional information can increase computational costs, especially for large-scale models and complex queries.
	\item Latency: Retrieval and processing can introduce latency, impacting the real-time performance of RAG systems.
	\item Complexity: Implementing and deploying RAG systems requires careful consideration of various factors, including data preparation, model selection, and system architecture.
\end{itemize}
\subsection{Interaction of Retrieval and Generation}
\begin{itemize}
	\item Aligning the goals of the retriever and generator is challenging.
	\item Optimizing the interaction between the two components requires careful design and tuning.
	\item The impact of various factors, such as metric selection and hyperparameter tuning, on RAG performance is still not fully understood.
	\item

\end{itemize}
\subsection{Long Context Generation:}

\begin{itemize}
	\item Context Length Limits: LLMs have limitations on the amount of context they can process at once.
	\item Information Loss: Long documents may be truncated or summarized, leading to loss of important information.
	\item Computational Cost: Processing long contexts can be computationally expensive.
\end{itemize}
\newpage
\section{Conclusion}
This chapter provided an overview of Retrieval-Augmented Generation (RAG), highlighting its significance in natural language processing. We discussed the key concepts of retrieval and generation, emphasizing the advantages of their combination in enhancing language model capabilities. The historical development of RAG was explored, along with its core components, including the retrieval mechanism, generation process, and augmentation techniques. Additionally, we examined various downstream tasks and evaluation targets, illustrating RAG's versatility in applications like open-domain question answering and fact verification. Overall, RAG represents a promising advancement in knowledge-intensive tasks, paving the way for further research and innovation in the field.
